---
title: "SI2019_clustering_exercise_draft_v1"
author: "pfr"
date: "August 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## --------------------------------------------------
##    PFR data prep exercises for clustering
## NOTE:
##   Use the W_mncntr matrix from the first code chunks 
##   from the dataprep SVD dimension reduciotn exercise
##
##  **** If W_mncntr does not exist then go back and run those chunks ***
## 
## --------------------------------------------------



##select numeric columns
First, SVD and PCA only work on numeric columns,
so we have to only keep the numeric columns


```{r get numeric columns, echo=TRUE}

# Get numeric columns only
col_classes = sapply(W_df,class)   #get column classes as a list
num_inds    = c(which(col_classes=='numeric'), which(col_classes=='integer'))
                 #get column number indices 
W_dfnum       = W_df[,num_inds]
dim(W_dfnum)

```


##Run Kmeans for several K values


```{r kmeans, echo=TRUE}

#use W_mncntr from dataprep_svd exercise

# Run kmeans for 20 values of K
kg=matrix(0,20,2)
for (i in 1:20){
  ktest=kmeans(W_mncntr,i,20,5);
  kg[i,1]=i;
  kg[i,2]=ktest$tot.withinss;  #save total within Sum Square Dist to center
}
plot(kg[,1],kg[,2],main='kmeans within cluster SS, weatherdata numerics')

#try
str(ktest)
#for example:
#  cluster has a cluster number assignments for each datum
#  centers are center points for each cluster
#  tot sum of squares
#  within sum of squares

```

##Now, let's combine the Kmeans cluster information with the 
## SVD reduced dataset

## NOTE:  The data matrix W_dfred should exist, 
##     
##  *** If W_dfred is not available, otherwise go back to SVD
##      exercise and rerun that   ***


```{r combine Kmeans cluster with SVD info, echo=TRUE}

#get Kmeans for 4 clusters 
k4             = kmeans(W_mncntr,4,20,5)
#get color scheme
col2use        = c('red','blue','black','yellow')
#get cluster assignment in colors
colassignments = col2use[k4$cluster]

W_proj = as.matrix(W_mncntr) %*% Wsvd$v[,c(1,2)] #    #Try different components?

plot(W_proj[,1],W_proj[,2],col=colassignments,main='data pts project to 1,2 SVD components, colored by kmeans')

# to plot center points, first project them into components
c3 = k4$centers%*% Wsvd$v[,1:3]          
points(c3[,1],c3[,2],pch=8,cex=2)           #Try different components?


```

## Now try coloring by class (using Y mean centered)


```{r plot with Y color}

#Y was created in SVD exercise, use it to select 2 colors

#get class assignment in colors
colassignments = col2use[Y]

plot(W_proj[,1],W_proj[,2],col=colassignments,main='data pts project to 1,2 SVD components, colored by class')




```
```{r lin model for prediciton}

linmodel_result = lm(Y~.,data=W_dfnum)   #Y is either 1 or 2, so let 1.5 be the cutoff of 'NO' rain prediction or 'YES'

#Note the 'Y~.'  is R's formula notation for Y=f( all the rest variables)
summary(Y)
```
```{r lin model for prediction cont}
#get model predictions for quick look at classification
Y_pred=linmodel_result$fitted.values  

#get the indices of predictions NO vs YES
Y_pred1_indices   =which(Y_pred<1.5)
Y_pred2_indices   =which(Y_pred>=1.5)

#set up No,Yes predictions
Y_pred_class                 =matrix(1,length(Y),1)  #default value of 1
Y_pred_class[Y_pred2_indices]=2                      #change value of Yes class

#show a confusion matrix 
table(Y,Y_pred_class)
```
```{r}
#Now color the predictions onto the 2 SVD dimensions

Yerr_ind  =which(Y!=Y_pred_class)
Yerr1_ind =intersect(which(Y==1),Yerr_ind)  #returns indices of class 1 errors
Yerr2_ind =intersect(which(Y==2),Yerr_ind)  #returns indices of class 2 errors

colassignments = col2use[Y_pred_class]

plot(W_proj[,1],W_proj[,2],col=colassignments,main='data pts project to 1,2 SVD components, colored by PREDICTED class')
```

```{r}

#Now plot the incorrect cases in different colors
Ycol_ind           =Y_pred_class
Ycol_ind[Yerr1_ind]=3     #set the color for errors, where true-value=1 (False Pos:blk)
Ycol_ind[Yerr2_ind]=4     #                    and true-value =2        (False Negs:yellow)
colassignments = col2use[Ycol_ind]



plot(W_proj[,1],W_proj[,2],col=colassignments,main='data pts project to 1,2 SVD components, Correct (R,Bl) and Errors (Y,Bk)') 

#Notice some of the incorrect class-1 cases are on the 'edges' of class 1 data in these dimensions, but other error cases
# are next to correctly predicted class-1 cases, which suggest that other dimensions have useful information

#How could we explore that?
#   go back to the original project and try other svd dimensions


#Note: ctrl-shift-k compiles report, or rmarkdown::render("SI2018_clustering_exercise_v0.Rmd") in #console
```

