---
title: "Data Exploration on Weather Data"
author: "SDSC Summer Institute"
modified: "2019-08-04"
date: "`r Sys.time()`"
output: html_document
---

## Preliminaries
### Read in weather data
```{r getData}
df <- readRDS(<<FILL-IN>>)
dim(df)
head(df$RISK_MM)
```

### Remove variable RISK_MM, which is the same as Rainfall for the next day
```{r  removeRiskMM}
df$RISK_MM <- <<FILL-IN>>
dim(df)
```

## Summary Statistics
### Check statistics and values in dataset
```{r stats}
<<FILL-IN>>(df)             # Structure of object
<<FILL-IN>>(df)             # Dimensions of object
<<FILL-IN>>(df)            # First few lines of data
<<FILL-IN>>(df)            # Last few lines of data
<<FILL-IN>>(df)           # Column headers
<<FILL-IN>>(df)         # Summary statistics
```

### Statistics on individual variables
```{r statsVar}

# Summary statistics
mean(df$MinTemp)
var(df$MinTemp)
sd(df$MinTemp)
summary(df$MinTemp)

# Other statistics
min(df$Date)                                # First date
max(df$Date)                                # Last date
sum(df$Rainfall)                            # Total rainfall amount for year
sum(df$Rainfall) / length(df$Rainfall)      # Average daily rainfall
```

### Remove rows with missing values
```{r na}

# Find number of places in dataset where there are missing values
sum(<<FILL-IN>>(df))

# Find number of samples (i.e., rows) without any missing value
sum(complete.cases(df))

# Remove rows with any missing value
dim(df)
df <- <<FILL-IN>>(df)
dim(df)
summary(df)
```

### Save the data as a CSV file (RISK_MM variable and NAs removed) to look at values
```{r saveCSV}
# NOTE: Order of factor levels is not preserved in csv files.
write.csv(df, <<FILL-IN>>, row.names=FALSE)
```

### Explore numerical variables
```{r numVars}
# Numerical variables (full set)
# numeric.vars <- c("MinTemp","MaxTemp","Rainfall","Evaporation","Sunshine","WindGustSpeed","WindSpeed9am","WindSpeed3pm","Humidity9am","Humidity3pm","Pressure9am","Pressure3pm","Cloud9am","Cloud3pm","Temp9am","Temp3pm")

# Subset of numerical variables.
numeric.vars <- c("Rainfall","Evaporation","Sunshine","WindSpeed9am","Humidity9am","Pressure9am","Cloud9am","Temp9am")

# Create DF with just numeric variables
df.num <- df[numeric.vars]
dim(<<FILL-IN>>)
head(df.num,3)
```

### Compute and plot pairwise correlation matrix
```{r corr}
<<FILL-IN>>(df.num, use="pairwise")
```

### Explore categorical variables
```{r ctgVars}

# Levels (categories) for WindGustDir
levels(df$WindGustDir)

# Number of occurrences for each WindGustDir category
<<FILL-IN>>(df$WindGustDir)

# Sorted number of occurrences for each level
sort(table(df$WindGustDir))

# Percentage for each WindGustDir category
round (table(df$WindGustDir) / sum(table(df$WindGustDir)) * 100, digits=2)

# Sorted percentages for WindGustDir categories
<<FILL-IN>> (round (table(df$WindGustDir) / sum(table(df$WindGustDir)) * 100, digits=2))
```

## Plots for visualizing data
### Histogram of Sunshine
```{r hist}
summary(df$Sunshine)
<<FILL-IN>>(df$Sunshine,breaks=10,col="lightblue",main="Histogram of Daily Sunshine",xlab="Sunshine (Hours)")
```

### Density plot of Sunshine
```{r density}
plot(density(df$Sunshine), main="Distribution of Daily Sunshine (Hours)")
plot(<<FILL-IN>>(df$Sunshine),ylim=c(0,0.2), main="Distribution of Daily Sunshine (Hours)")
```

### Bar plot showing frequency of wind gust directions
```{r barplot}
table(df$WindGustDir)
plot(df$<<FILL-IN>>,col="lightblue",main="Distribution of Wind Gust Direction",xlab="Wind Gust Direction",ylab="Count")
```

### Scatter plot of Evaporation vs. Temperature
```{r scatterplot}
cor(df$Evaporation,df$Temp9am)
plot(df$Evaporation,df$<<FILL-IN>>,col="blue",main="Evaporation vs. Temperature",xlab="Evaporation (mm)",ylab="Temperature")
```

### Scatter plot of Evaporation vs. Temperature with regression line
```{r scatterplot2}
plot(df$Evaporation,df$Temp9am,col="blue",main="Evaporation vs. Temperature",xlab="Evaporation (mm)",ylab="Temp9am (C)")
abline(lm(df$Temp9am ~ df$Evaporation), col="red")
```

### Line plot of daily high temperature
```{r lineplot}
# Get summary statistics
summary(df$MaxTemp)        

# Using base R plot command
plot(df$Date,df$<<FILL-IN>>,type='l',col="blue",main="Daily High Temperature",xlab="Date",ylab="MaxTemp (C)")

```

### Box plot of atmospheric pressure with respect to RainToday
```{r boxplot}
# install.packages("ggplot2")
library(ggplot2)
<<FILL-IN>>(df, aes(x=RainToday, y=Pressure9am, fill=RainToday)) + geom_boxplot() +
  ggtitle("Atmospheric Pressure wrt Rain") 
```

### Box plot of MaxTemp by month
```{r boxplot2}
# install.packages("lubridate")
library(lubridate)   # R package for handling dates
ggplot(df, aes(x=as.factor(<<FILL-IN>>(Date)), y=MaxTemp)) + geom_boxplot(fill="lightblue") + 
  xlab("Month") + ylab("MaxTemp (Celsius)") + ggtitle("High Temperature by Month")
```

### Violin plot of MaxTemp by month
```{r violinplot}
ggplot(df, aes(x=as.factor(month(Date)), y=MaxTemp)) + geom_<<FILL-IN>>(fill="lightblue") + 
  xlab("Month") + ylab("MaxTemp (Celsius)") + ggtitle("High Temperature by Month")
```

### Plot correlation matrix
```{r corrplot}
# install.packages("GGally")
library(GGally)        # R package that extends ggplot2  
cor(df.num, use = "pairwise")
<<FILL-IN>>(df.num, label=TRUE, label_round=2)  # ggcorr uses cor to calculate correlations
```

### Plot pairwise correlation
```{r pairsplot}
<<FILL-IN>>(df[c("Sunshine","Pressure9am","Humidity9am","Temp9am")])  # ggpairs is from GGally package
```